Review: Crop adaptation through introgression from wild relatives

### 2017-04-06 paper.tex
Update paper.tex

In an attempt to categorize the varieties of Asian rice, \cite{khush2003classifying} performed isozyme polymorphism analysis (11 enzymes, 20 loci, 79 alleles) on 25,519 rice varieties.
The vast majority of the rice varieties fell into two groups; indica (73.45\%) and japonicas (23.02\%).
Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization of ancient japonica with new wild rice populations in south and southeast Asia \cite{Huang2012, londo2006phylogeography}.
However,about 1.26\% belonged to four other groups (and 2.27\% were indeterminate).
Some of the varieties from these other four groups most likely arose from hybrids of cultivated rice and \emph{O. rufipogon}.
Introgression of alleles between cultivated rices, as well as between cultivated and wild rices, is common \cite{oka2012origin, second1982origin, zhao2010genomic}\, and natural introgression from wild to domesticate is suspected (\cite{zhao2010genomic} calls for research into this possibility).
These rices often have blight resistance gene, likley imparted from the wild relative parent.
Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) are known to have been introgressed from wild relatives into \emph{O. sativa} by researchers \cite{brar1997alien, khush1974inheritance}.
Some aus and rayada varieties retain characteristics of the putative wild parent.
Aus exhibits a sprawling growth pattern and easy-threshing grain.
Rayada is adapted to areas with longer periods of flooding, with heightened seed dormancy and photoperiod sensitivity (permitting harvest during times without standing water).
Aswina variety, adapted to growth in deepwater conditions with a short period of flooding, seems to be the result of hybridization between O. sativa and deepwater-adapted O. rufipogon populations.

@@

Domesticated and wild barleys belong to the same species, \emph{Hordeum vulgare}, and are biologically capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) is believed to have been domesticated at least twice, once from wild \emph{subsp. spontaneum} in the Fertile Crescent and once from subsp. \emph{spontaneum} var. \emph{agriocrithon} in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\, but many of the details of barley domestication are still disputed.

@@

The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
Pre-Columbian H. annuus distribution spanned much of the Great Plains, from what is now north-central Texas up to and through Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
Domesticated sunflower has long lived in sympatry with wild relatives like \emph{H. petiolaris} and \emph{H. bolanderi} and forms stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
Many wild sunflowers are locally-adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}.
However, the most striking example of adaptive introgression within Helianthus is that of the cucumberleaf sunflower, H. debilis ssp. cucumerifolius.
Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.
Introgressive hybridization imparted locally-adapted alleles from H. debilis to H. annuus via introgressive hybridization. 
These introgressed hybrids formed a new lineage of sunflower (H. annuus ssp. texanus, H. a. texanus hereafter) which displays H. debilis-like traits adaptive to south-central Texas climate and ecology.
These adaptive debilis-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than H. annuus (as measured by higher seed production)\cite{whitney2006adaptive}.
Although H. annuus and H. a. texanus are interfertile, H. a. texanus displays persistent phenotypic differences from H. annuus.

@@

Recent investigations into the history of Helianthus introgression have implemented genomic methods.
\cite{Baute2015} analyzed transcriptome sequence variation on cultivated and wild H. annuus, H. petiolaris, and H. argophyllus.
Using STRUCTURE, these authors found that introgressions from wild relatives can be found on every chromosome in at least one modern line, covering over 10\% of the genome.
Of particular note is the modern line RHA 274, a modern line which was bred with H. a. texanus in the 1970s to restore a branching plant body architecture, which allows the plant to produce pollen for a longer period of time, increasing seed production.
RHA 274 has several large introgression from H. a. texanus, including one at the site of HaGNAT, the domestication gene associated with branching.
These introgressed regions are not found in the non-branching lines Sunrise and VNIIMK8931, further suggesting that the H. a. texanus introgressed regions are causative.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=17.35cm]{map.png}
	\caption{Map of the domestication centers of four crops and the distribution of their interfertile wild relatives, overlaying average annual precipitation.}
	\label{boxplot:map}
\end{figure}

>>>>

In an attempt to categorize the varieties of Asian rice, \cite{khush2003classifying} performed isozyme polymorphism analysis (11 enzymes, 20 loci, 79 alleles) on 25,519 rice varieties.
The vast majority of the rice varieties fell into two groups; indica (73.45\%) and japonicas (23.02\%).
Japonica rice cultivars were likely domesticated first from wild rice populations in southern China, and indica cultivars later developed through hybridization of ancient japonica with new wild rice populations in south and southeast Asia \cite{Huang2012, londo2006phylogeography}.
However,about 1.26\% belonged to four other groups (and 2.27\% were indeterminate).
Some of the varieties from these other four groups most likely arose from hybrids of cultivated rice and \emph{O. rufipogon}.
Introgression of alleles between cultivated rices, as well as between cultivated and wild rices, is common \cite{oka2012origin, second1982origin, zhao2010genomic}\, and natural introgression from wild to domesticate is suspected (\cite{zhao2010genomic} calls for research into this possibility).
These rices often have blight resistance gene, likley imparted from the wild relative parent.
Several resistance genes (grassy stunt virus, bacterial blight, brown planthopper, blast) are known to have been introgressed from wild relatives into \emph{O. sativa} by researchers \cite{brar1997alien, khush1974inheritance}.
Some aus and rayada varieties retain characteristics of the putative wild parent.
Aus exhibits a sprawling growth pattern and easy-threshing grain.
Rayada is adapted to areas with longer periods of flooding, with heightened seed dormancy and photoperiod sensitivity (permitting harvest during times without standing water).
Aswina variety, adapted to growth in deepwater conditions with a short period of flooding, seems to be the result of hybridization between \emph{O. sativa} and deepwater-adapted \emph{O. rufipogon} populations.

@@

Domesticated and wild barleys belong to the same species, \emph{Hordeum vulgare}, and are biologically capable of producing viable offpsring via hybridizaion \cite{von1995ecographical}.
Barley (\emph{Hordeum vulgare} subsp. \emph{vulgare}) is believed to have been domesticated at least twice, once from wild subsp. \emph{spontaneum} in the Fertile Crescent and once from subsp. \emph{spontaneum} var. \emph{agriocrithon} in Tibet roughly 10,000 years ago \cite{takahashi1955origin, badr2000origin, oka2012origin, azhaguvel2007phylogenetic, haberer2015barley}\, but many of the details of barley domestication are still disputed.

@@

The common sunflower (\emph{Helianthus annuus}) shows evidence of domestication in eastern United States \cite{harter2004origin, wills2006chloroplast}\, with additional evidence of a possible second origin of domestication in Mexico \cite{lentz2008sunflower}\.
Pre-Columbian \emph{H. annuus} distribution spanned much of the Great Plains, from what is now north-central Texas up to and through Montana and North Dakota (see figure 1 of \cite{whitney2010adaptive}).
Domesticated sunflower has long lived in sympatry with wild relatives like \emph{H. petiolaris} and \emph{H. bolanderi} and forms stable hybrid populations \cite{schwarzbach2002likely, rieseberg1988molecular, welch2002patterns}\.
Many wild sunflowers are locally-adapted, and weedy hybrid populations share these adaptations \cite{kane2008genetics}.
However, the most striking example of adaptive introgression within \emph{Helianthus} is that of the cucumberleaf sunflower, \emph{H. debilis} ssp. \emph{cucumerifolius}.
Cucumberleaf sunflower is endemic to south-central Texas, and exhibits several adaptations to the region.
Introgressive hybridization imparted locally-adapted alleles from \emph{H. debilis} to \emph{H. annuus} via introgressive hybridization \cite{heiser1951hybridization}. 
These introgressed hybrids formed a new lineage of sunflower (\emph{H. annuus} ssp. \emph{texanus}, \emph{H. a. texanus} hereafter) which displays \emph{H. debilis}-like traits adaptive to south-central Texas climate and ecology.
These adaptive debilis-like traits include resistance to herbivorous pests and an increased branching plant architecture, as well as higher overall fitness than \emph{H. annuus} (as measured by higher seed production)\cite{whitney2006adaptive}.
Although H. annuus and \emph{H. a. texanus} are interfertile, \emph{H. a. texanus} displays persistent phenotypic differences from \emph{H. annuus} \cite{rieseberg2007hybridization}.

@@

Recent investigations into the history of \emph{Helianthus} introgression have implemented genomic methods.
\cite{Baute2015} analyzed transcriptome sequence variation on cultivated and wild \emph{H. annuus}, \emph{H. petiolaris}, and \emph{H. argophyllus}.
Using STRUCTURE, these authors found that introgressions from wild relatives can be found on every chromosome in at least one modern line, covering over 10\% of the genome.
Of particular note is the modern line RHA 274, a modern line which was bred with H. a. texanus in the 1970s to restore a branching plant body architecture, which allows the plant to produce pollen for a longer period of time, increasing seed production.
RHA 274 has several large introgression from \emph{H. a. texanus}, including one at the site of HaGNAT, the domestication gene associated with branching.
These introgressed regions are not found in the non-branching lines Sunrise and VNIIMK8931, further suggesting that the \emph{H. a. texanus} introgressed regions are causative.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=17.35cm]{figure_outline.png}
	\caption{Map of the domestication centers of four crops and the distribution of their interfertile wild relatives, overlaying average annual precipitation.}
	\label{boxplot:map}
\end{figure}
